##Loading and preprocessing the data##

```{r}
activity_DF <- read.csv("activity.csv")
##activity_DF <- read.csv("./Coursera/activity.csv")
```

# Transform the date attribute to an actual date format
```{r}
activity_DF$date <- as.POSIXct(activity_DF$date, format="%Y-%m-%d")
```

# Calculate the weekdays from the date attribute
```{r}
activity_DF <- data.frame(date=activity_DF$date, 
                          weekday=tolower(weekdays(activity_DF$date)), 
                          steps=activity_DF$steps, 
                          interval=activity_DF$interval)
```

# Calculate the day type (weekend or weekday)
```{r}
activity_DF <- cbind(activity_DF, 
                     daytype=ifelse(activity_DF$weekday == "saturday" | 
                                      activity_DF$weekday == "sunday", "weekend", 
                                    "weekday"))
```

# Create the final data.frame
```{r}
activity <- data.frame(date=activity_DF$date, 
                       weekday=activity_DF$weekday, 
                       daytype=activity_DF$daytype, 
                       interval=activity_DF$interval,
                       steps=activity_DF$steps)
rm(activity_DF)
```
# Aggregating(sum) the dataset by day#
```{r}
Means_Steps <- aggregate(steps ~ date, activity, FUN = sum, na.rm = TRUE)
```

# Make a histogram for total steps per day#
```{r}
hist(Means_Steps$steps, 
     main=" ",
     breaks=seq(from=0, to=25000, by=2500),
     col="purple",
     xlab="Total Number of Steps Taken per day")
```

##Calculate and report the mean and median of the total number of steps taken per day#
```{r}
mean(Means_Steps$steps)
```

```{r}
median(Means_Steps$steps)
```
##What is the average daily activity pattern?#
#Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
Avg_Steps <- aggregate(activity$steps, 
          by=list(activity$interval), 
          FUN=mean, 
          na.rm=TRUE)

names(Avg_Steps) <- c("interval", "mean")
plot(Avg_Steps, type = "l")
```

##Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

## We find the position of the maximum mean
```{r}
maxmean_pos <- which(Avg_Steps$mean == max(Avg_Steps$mean))
```
## We lookup the value of interval at this position
```{r}
Avg_Steps[maxmean_pos, 1]
```

#Imputing missing values

##Calculate and report the total number of missing values in the dataset
#         (i.e. the total number of rows with NAs)
```{r}
sum(is.na(activity))
```

## Find the NA positions
```{r}
na_values <- which(is.na(activity$steps))
```
## Create a vector of means
```{r}
mean_values <- rep(mean(activity$steps, na.rm=TRUE), times=length(na_values))
```

```{r}
activity[na_values, "steps"] <- mean_values
```

```{r}
head(activity)
```
##4. Make a histogram of the total number of steps taken each day and calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

### Compute the total number of steps each day (NA values removed)
```{r}
sum_activity <- aggregate(activity$steps, by=list(activity$date), FUN=sum)
```
## Rename the attributes
```{r}
names(sum_activity) <- c("date", "total")
```
## Compute the histogram of the total number of steps each day
```{r}
hist(sum_activity$total, 
     breaks=seq(from=0, to=25000, by=2500),
     col="pink", 
     xlab="Total number of steps", 
     ylim=c(0, 30), 
     main="Total no. of steps each day- Histogram (NA replaced by mean value)")
```


##  Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
## "daytype" variable is already created in the activity dataset
```{r}
head(activity)
```
```{r}
rm(sum_activity)
```
####    Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).


### Load the lattice graphical library
```{r}
library(lattice)
```
### Compute the average number of steps taken, averaged across all daytype variable
```{r}
mean_DF <- aggregate(activity$steps, 
                       by=list(activity$daytype, 
                               activity$weekday, activity$interval), mean)
```
### Rename the attributes
```{r}
names(mean_DF) <- c("daytype", "weekday", "interval", "mean")
# Compute the time series plot
xyplot(mean ~ interval | daytype, mean_DF, 
       type="l", 
       lwd=1, 
       xlab="Interval", 
       ylab="Number of steps", 
       layout=c(1,2))
```